\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-10-09 chapters/conclusion.tex
Self-plagiarism check

\chapter{Conclusion \& recommendations}

@@

\subsection{Developed \acs{erp} decoders}
We introduced a covariance estimator using adaptive shrinkage (STBF-shrunk) and an estimator
exploiting prior knowledge about the spatiotemporal nature of the EEG signal
(STBF-struct).
We compared these estimators with the original formulation of the
spatiotemporal (STBF-emp)
beamformer and a state-of-the-art Riemannian Geometry method (XDAWN+RG) in an off-line P3 detection task on
an existing dataset.
Our results show that the structured estimator results in higher accuracy in
general, and specifically when training data are sparsely available.
Results can be computed faster and with
substantially less memory usage.
Since these algorithms are not paradigm-specific, the conclusions can be
generalized to other ERP-based BCI settings.
These results have been published in~\textcite{VanDenKerchove2022}.

Next, \acf{bttda} was introduced as a tensor-based decoder that better captures the
multidimensional nature of \ac{erp} data.
By preserving the structure of neural data, \ac{bttda} can cope with noise and other challenges
arising from gaze-independent \acp{bci}, yielding robust results under certain conditions.
Results show that \ac{bttda} and its special sum-of-rank-one structured case
improved over \ac{hoda} and can reach state-of-the-art decoding performance for
\acp{erp}.
This work is submitted as~\textcite{VanDenKerchove2024}.

>>>>

\chapter{Conclusion \& recommendations}
\emph{The first paragraph of \cref{sec:conclusion/contrib/decoders} was published as
part of \textcite{VanDenKerchove2022}.}

@@

\subsection{Developed \acs{erp} decoders}
\label{sec:conclusion/contrib/decoders}
We introduced a covariance estimator using adaptive shrinkage (STBF-shrunk) and an estimator
exploiting prior knowledge about the spatiotemporal nature of the EEG signal
(STBF-struct).
We compared these estimators with the original formulation of the
spatiotemporal (STBF-emp)
beamformer and a state-of-the-art Riemannian Geometry method (XDAWN+RG) in an off-line P3 detection task on
an existing dataset.
Our results show that the structured estimator results in higher accuracy in
general, and specifically when training data are sparsely available.
Results can be computed faster and with
substantially less memory usage.
Since these algorithms are not paradigm-specific, the conclusions can be
generalized to other ERP-based BCI settings.
These results have been published in~\textcite{VanDenKerchove2022}.

Next, \acf{bttda} was introduced as a tensor-based decoder that better captures the
multidimensional nature of \ac{erp} data.
By preserving the structure of neural data, \ac{bttda} can cope with noise and other challenges
arising from gaze-independent \acp{bci}, yielding robust results under certain conditions.
Results show that \ac{bttda} and its special sum-of-rank-one structured case
improved over \ac{hoda} and can reach state-of-the-art decoding performance for
\acp{erp}.
This work is submitted as~\textcite{VanDenKerchove2024a}.

### 2024-10-09 chapters/contribution.tex
Self-plagiarism check

\chapter*{Personal contribution}
Arne Van Den Kerchove conceptualized the works and performed the analyses in
this thesis, under supervision and in consultation with prof. Hakim Si-Mohammed,
prof. Marc Van Hulle and prof. François Cabestaing.
Dr. Arno Libert and Dr. Benjamin Wittevrongel consulted on the analyses performed in
\cref{sec:stbf-struct}
Data collection was performed by Arne Van Den Kerchove for the dataset
presented in \cref{sec:covert-align} and by Arne Van Den Kerchove and
Juliette Meunier for the dataset presented in \cref{sec:patients}.
Participant recruitment at TRAINM was performed by Alixe Willemssens under
supervision of dr. Edward Schietekatte, at Fondation Partage et Vie by Marie de
Moura under supervision of prof. Etienne Allart and at the UZ Leuven Neuromuscular
Reference Center under supervision of prof. Kristl Claeys and prof. Philip Van Damme.
Prof. Etienne Allart provided clinical consultation.

>>>>

\chapter*{Personal contribution}
Arne Van Den Kerchove conceptualized the works and performed the analyses
presented in
this thesis, under supervision and in consultation with prof. Hakim Si-Mohammed,
prof. Marc Van Hulle and prof. François Cabestaing.
Dr. Arno Libert and Dr. Benjamin Wittevrongel consulted on the analyses performed in
\cref{sec:stbf-struct}
Data collection was performed by Arne Van Den Kerchove for the dataset
presented in \cref{sec:covert-align} and by Arne Van Den Kerchove and
Juliette Meunier for the dataset presented in \cref{sec:patients}.
Participant recruitment at TRAINM was performed by Alixe Willemssens under
supervision of dr. Edward Schietekatte, at Fondation Partage et Vie by Marie de
Moura under supervision of prof. Etienne Allart and at the UZ Leuven Neuromuscular
Reference Center under supervision of prof. Kristl Claeys and prof. Philip Van Damme.
Prof. Etienne Allart provided clinical consultation.

### 2024-10-09 chapters/covert_align.tex
Self-plagiarism check

\chapter{Compensating \acs{erp} jitter for gaze-independence}
\label{sec:covert-align}
\emph{This chapter has been published as part of~\textcite{VanDenKerchove2024}.}

@@

\subsection{BCI decoding performance}
\label{sec:block_accuracy}

We evaluated the BCI decoding performance in a single-trial classification
experiment, as well as in a target selection experiment reflecting BCI
operation.

\Cref{fig:single_trial_roc_auc_diff} shows a comparison of areas under the
Receiver Operating Characteristic curve (ROC-AUC) for all pairs of \ac{tlda}, \ac{cble},
and \ac{wcble} for single-trial classification to investigate the contributions of
\ac{cble} and \ac{wcble} relative to their first-stage classifier \ac{tlda}.
For this evaluation, epochs were rejected when the peak-to-peak amplitude
exceeded 800$\mu V$, and for the CVSA-ERP dataset, if the user's gaze differed by more than
10 degrees of visual angle from the fixation crosshair.
\begin{figure}
  \bigskip
	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:single_trial_roc_auc_diff}
		\includegraphics[width=\linewidth]{figures/covert_align/figure4a.pdf}
	\end{subfigure}

@@

\Cref{fig:block_evaluation} shows the cross-validated BCI selection
accuracy on the BNCI2014-009 and the CVSA-ERP datasets for all investigated
decoders.
Accuracy was determined by, for each block, selecting the character with the
highest (stage-two if applicable) classifier score and comparing it to the cued target.
Significance was calculated using one-sided Wilcoxon rank-sum tests ($p=0.05$)
corrected for FDR over decoders.
For this evaluation, no epochs were rejected to keep the trial-based structure of BCI
operation intact.
For all datasets and \ac{vsa} conditions, \ac{cble} scores approximately on par with \ac{tlda}.
Yet, \ac{wcble} yields an improved decoding accuracy for covert \ac{vsa} in both
datasets, which is greatest for smaller numbers of repetitions and
decreases as the number of repetitions increases.
This covert \ac{vsa} accuracy increase over \ac{tlda} is significant in thdeep q learning tutoriale BNCI2014-009 dataset
for 1 and 3 repetitions, and in CVSA-ERP for 1-5 and 10 repetitions.
Furthermore, while we reported a relative decrease in single-trial ROC-AUC for
\ac{wcble} in overt \ac{vsa}, this does not seem to result in a consistent decrease in target
selection accuracy.
No significant increase of \ac{wcble} over other methods was found in split \ac{vsa}.
While Riemannian methods are significantly outperformed by \ac{tlda}, \ac{cble}, and \ac{wcble}
in the BNCI2014-009 dataset, they perform approximately on par with \ac{tlda} and
\ac{wcble} in CVSA-ERP.

>>>>

\chapter{Compensating \acs{erp} jitter for gaze-independence}
\label{sec:covert-align}
\emph{This chapter was published as part of~\textcite{VanDenKerchove2024}.}

@@

\subsection{\Acs{bci} decoding performance}
\label{sec:block_accuracy}

We evaluated the \ac{bci} decoding performance in a single-trial classification
experiment, as well as in a target selection experiment reflecting \ac{bci}
operation.

\Cref{fig:single_trial_roc_auc_diff} shows a comparison of \ac{rocauc}
for all pairs of \ac{tlda}, \ac{cble},
and \ac{wcble} for single-trial classification to investigate the contributions of
\ac{cble} and \ac{wcble} relative to their first-stage classifier \ac{tlda}.
For this evaluation, epochs were rejected when the peak-to-peak amplitude
exceeded 800$\mu V$, and for the CVSA-ERP dataset, if the user's gaze differed by more than
10 degrees of visual angle from the fixation crosshair.
\begin{figure}
  \bigskip
	\begin{subfigure}{\linewidth}
		\caption{}
		\label{fig:single_trial_roc_auc_diff}
		\includegraphics[width=\linewidth]{figures/covert_align/figure4a.pdf}
	\end{subfigure}

@@

\Cref{fig:block_evaluation} shows the cross-validated \ac{bci} selection
accuracy on the BNCI2014-009 and the CVSA-ERP datasets for all investigated
decoders.
Accuracy was determined by, for each block, selecting the character with the
highest (stage-two if applicable) classifier score and comparing it to the cued target.
Significance was calculated using one-sided Wilcoxon rank-sum tests ($p=0.05$)
corrected for FDR over decoders.
For this evaluation, no epochs were rejected to keep the trial-based structure
of \ac{bci} operation intact.
For all datasets and \ac{vsa} conditions, \ac{cble} scores approximately on par with \ac{tlda}.
Yet, \ac{wcble} yields an improved decoding accuracy for covert \ac{vsa} in both
datasets, which is greatest for smaller numbers of repetitions and
decreases as the number of repetitions increases.
This covert \ac{vsa} accuracy increase over \ac{tlda} is significant in thdeep q learning tutoriale BNCI2014-009 dataset
for 1 and 3 repetitions, and in CVSA-ERP for 1-5 and 10 repetitions.
Furthermore, while we reported a relative decrease in single-trial \ac{rocauc} for
\ac{wcble} in overt \ac{vsa}, this does not seem to result in a consistent decrease in target
selection accuracy.
No significant increase of \ac{wcble} over other methods was found in split \ac{vsa}.
While Riemannian methods are significantly outperformed by \ac{tlda}, \ac{cble}, and \ac{wcble}
in the BNCI2014-009 dataset, they perform approximately on par with \ac{tlda} and
\ac{wcble} in CVSA-ERP.

### 2024-10-09 chapters/stbf_struct.tex
Self-plagiarism check

	%Besides the recording method, \acp{bci} differ in the communication paradigms used
	%for communication~\cite{Abiri2019}.
	%A popular class of \ac{bci} paradigms relies on the evocation
	%of Event-Related Potentials (\acp{erp}) in the brain in response to visual, auditory, or tactile stimulation, given their low decoding cost and generally short
	%calibration time before usage~\cite{Gao2014, Kapgate2015}.
	%The study we report on focuses on the visual P300 oddball \ac{erp} in response to a
	%rare but attended visual stimulus.
	%The decoder detects whether this \ac{erp} is present to determine which stimulus
	%the user attended.
	%The P300 paradigm has been used extensively in \ac{bci} development and is easy to
	%set up~\cite{Farwell1988, Sellers2006, Barachant2014, Philip2020}.

  There are multiple state-of-the-art P300 classification methods, like
  \acp{svm}~\cite{Tayeb2014}, deep
	learning models~\cite{Vareka2020,Borra2020}, and Riemannian Geometry
	classifiers~\cite{Barachant2014}.
	While these models often return a high classification accuracy, there is a need
	for lightweight models -- lightweight models lead to fast off-line analyses and can be transferred to consumer-grade hardware.
	When moving towards plug-and-play solutions, \ac{bci} calibration sessions should be short and model training times low.
  The \ac{stbf}~\cite{VanVliet2015, Wittevrongel2016} belongs to
	this class of \ac{erp} decoding models as it achieves state-of-the-art performance and is fast to train.
	Earlier work has shown that it is
	possible to apply the spatiotemporal beamformer to multiple time-locked visual
  \ac{bci} paradigms, including the P300 oddball paradigm, \ac{ssvep},
  \ac{cvep}~\cite{Wittevrongel2017a}, \ac{mvep}~\cite{Libert2021}.

@@

	\subsection{Dataset}
	We use the dataset from~\textcite{Wittevrongel2016}, containing P300 oddball \ac{eeg}
	recordings of 21 healthy subjects since it is a high-quality dataset with a high
	number (32) of electrodes and concurrently recorded EOG responses for ocular artifact rejection.
	Nine targets were arranged on a monitor before the subject during an
	experimental session.
	The subject was asked to pay attention to a cued target for a block
	of stimulations.
	The stimulations in a block are organized in 15 separate subsequent trials.
	A trial is defined as 9 stimulations in which each target is flashed
	precisely once per trial.
	Each target was cued four times, resulting in a dataset consisting of 36 blocks
	(4860 stimulations) per subject.
	Each stimulation will correspond to a single epoch in the preprocessed dataset.
	See~\textcite{Wittevrongel2016} for a complete description of the dataset and the recording procedure.

@@

	%	the central and parietal regions, where the P300 \ac{erp} component is present.

@@

	%	that state that, in addition to the P300, the early N1 and P2 \ac{erp} components

@@

	\section{Conclusion}
	While it is possible to regularize the spatiotemporal \ac{lcmv} beamformer
	classifier for \ac{erp} detection with other methods such as by employing feature selection,
	by adding regularizing penalties to the cost
	function beamforming problem, or by crafting a cleaner activation pattern, our work focused on
	estimation methods for the spatiotemporal covariance.
	We introduced a covariance estimator using adaptive shrinkage and an
	estimator exploiting prior knowledge about the spatiotemporal nature of the \ac{eeg}
	signal.
	We compared these estimators with the original spatiotemporal
	beamformer and a state-of-the-art method in an off-line P300 detection task.
	Our results show that the structured estimator performs better when training data are sparsely available and that it can be computed faster and with substantially less memory usage.
	Since these algorithms are not paradigm-specific,  the conclusions can be generalized to
	other \ac{erp}-based \ac{bci} settings.

>>>>

	%Besides the recording method, \acp{bci} differ in the communication paradigms used
	%for communication~\cite{Abiri2019}.
	%A popular class of \ac{bci} paradigms relies on the evocation
	%of Event-Related Potentials (\acp{erp}) in the brain in response to visual, auditory, or tactile stimulation, given their low decoding cost and generally short
	%calibration time before usage~\cite{Gao2014, Kapgate2015}.
	%The study we report on focuses on the visual P3 oddball \ac{erp} in response to a
	%rare but attended visual stimulus.
	%The decoder detects whether this \ac{erp} is present to determine which stimulus
	%the user attended.
	%The P3 paradigm has been used extensively in \ac{bci} development and is easy to
	%set up~\cite{Farwell1988, Sellers2006, Barachant2014, Philip2020}.

  There are multiple state-of-the-art P3 classification methods, like
  \acp{svm}~\cite{Tayeb2014}, deep
	learning models~\cite{Vareka2020,Borra2020}, and Riemannian Geometry
	classifiers~\cite{Barachant2014}.
	While these models often return a high classification accuracy, there is a need
	for lightweight models -- lightweight models lead to fast off-line analyses and can be transferred to consumer-grade hardware.
	When moving towards plug-and-play solutions, \ac{bci} calibration sessions should be short and model training times low.
  The \ac{stbf}~\cite{VanVliet2015, Wittevrongel2016} belongs to
	this class of \ac{erp} decoding models as it achieves state-of-the-art performance and is fast to train.
	Earlier work has shown that it is
	possible to apply the spatiotemporal beamformer to multiple time-locked visual
  \ac{bci} paradigms, including the P3 oddball paradigm, \ac{ssvep},
  \ac{cvep}~\cite{Wittevrongel2017a}, \ac{mvep}~\cite{Libert2021}.

@@

	\subsection{Dataset}
	We use the dataset from~\textcite{Wittevrongel2016}, containing P3 oddball \ac{eeg}
	recordings of 21 healthy subjects since it is a high-quality dataset with a high
	number (32) of electrodes and concurrently recorded EOG responses for ocular artifact rejection.
	Nine targets were arranged on a monitor before the subject during an
	experimental session.
	The subject was asked to pay attention to a cued target for a block
	of stimulations.
	The stimulations in a block are organized in 15 separate subsequent trials.
	A trial is defined as 9 stimulations in which each target is flashed
	precisely once per trial.
	Each target was cued four times, resulting in a dataset consisting of 36 blocks
	(4860 stimulations) per subject.
	Each stimulation will correspond to a single epoch in the preprocessed dataset.
	See~\textcite{Wittevrongel2016} for a complete description of the dataset and the recording procedure.

@@

	%	the central and parietal regions, where the P3 \ac{erp} component is present.

@@

	%	that state that, in addition to the P3, the early N1 and P2 \ac{erp} components

@@

	\section{Conclusion}
	While it is possible to regularize the spatiotemporal \ac{lcmv} beamformer
	classifier for \ac{erp} detection with other methods such as by employing feature selection,
	by adding regularizing penalties to the cost
	function beamforming problem, or by crafting a cleaner activation pattern, our work focused on
	estimation methods for the spatiotemporal covariance.
	We introduced a covariance estimator using adaptive shrinkage and an
	estimator exploiting prior knowledge about the spatiotemporal nature of the \ac{eeg}
	signal.
	We compared these estimators with the original spatiotemporal
	beamformer and a state-of-the-art method in an off-line P3 detection task.
	Our results show that the structured estimator performs better when training data are sparsely available and that it can be computed faster and with substantially less memory usage.
	Since these algorithms are not paradigm-specific,  the conclusions can be generalized to
	other \ac{erp}-based \ac{bci} settings.

### 2024-10-09 chapters/wcble.tex
Self-plagiarism check

\chapter{\Acs{erp} latency estimation and alignment}
\label{sec:wcble}
\emph{Sections~\ref{sec:wcble/literature/contrib} and~\ref{sec:wcble/methods} were published
as part of~\textcite{VanDenKerchove2024}.}

@@

The \ac{wcble} algorithm is presented in \cref{sec:wcble/app/wcble-alg}.
Its training phase is visualized in~\cref{fig:diagram-train}.
The initial training epochs $\{\mat{X}_n^{(1)}\}_{n=1}^N$ are set to $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N$.
At every iteration, classifier $\mathcal{C}$ is trained like in \ac{cble}:
\begin{equation}
  \theta^{(i)} =
  \mathrm{train}_\mathcal{C}(\{\mat{X}^{(i-1)}_n[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
\end{equation}
Next, latency $s_{l_n}^{(i)}$ is determined for every epoch $\mat{X}^{(i)}$ corresponding
to its class label $l_n$ using \cref{eq:prob-latency}.
Finally, the training epochs $\mat{X}^{(i+1)}$ for the next iteration are determined by aligning
each original training epoch to the latency $s_{l_n}^{(i)}$ corresponding to its respective class
label.
\begin{equation}
  \mat{X}^{(i+1)}_n = \mathrm{align}(\mat{X}_n^\mathrm{train}, s_{l_n}^{(i)}) \quad \forall n=1,\ldots,N
\end{equation}
Aligning is performed by shifting and zero-padding the signal to the right if
the latency is negative relative to the time window onset, and to the left if
positive, by the difference between the latency and the window onset.
\fullpagefig{%
  \includegraphics[width=\textwidth]{figures/wcble/figure1.pdf}
}{%
  Schematic representation of the \acs{wcble} training phase.
}{%
  Schematic representation of the \ac{wcble} training phase.
  (1) The first-stage spatiotemporal binary classifier is trained on a set of
  epochs.
  (2) It is then applied to time-shifted copies of the epochs to obtain scores
  and class probabilities over time.
  (3) The medians of these probability distributions are assumed as the class
  latencies.
  (4) The epochs are aligned to their corresponding class latencies by
  shifting in time such that all latencies fall at the same moment.
  (5) The spatiotemporal classifier is then retrained on the aligned epochs
  for the next iteration.
  (6) After the iterative process halts, the scores and latencies obtained
  from the last iteration are used to train the second-stage
  classifier.
}{fig:diagram-train}
The process halts after a fixed number of iterations or when the estimated set
of latencies has been encountered before, indicating it ended up in a loop.
In the end, the procedure should result in enhanced classifier parameters $\theta^*$,
closer to those when there would be no jitter between epochs.
Note that using the median approach for robust latency estimation results in a
smoother yet longer convergence process compared to the maximum/minimum approach.
We can then apply the classifier with enhanced parameters $\theta^*$ in a \ac{cble}
manner to unseen epochs as illustrated in \cref{fig:diagram-eval} to obtain a
vector of scores over time as in \cref{sec:wcble/methods/cble} and the estimated latencies as
in~\cref{sec:robust-latency}.
\fullpagefig{
  \includegraphics[width=\textwidth]{figures/wcble/figure2.pdf}

}{%
  Schematic representation of the \acs{wcble} test phase.
}{%
  Schematic representation of the Woody Classifier-Based Latency
  Estimation test phase. (1) The
	first-stage spatiotemporal binary classifier obtained from the training phase is
	applied to time-shifted copies of the epochs to obtain scores and class probabilities
	over time. (2) The medians of these probability distributions are assumed
	as the new class latencies. (3) The scores and class latencies are input to the
  trained second-stage classifier, which predicts the label of the epochs.
}{fig:diagram-eval}
\section{Results}
\label{sec:wcble/results}
Since it is hard to obtain ground truth measures of \ac{erp} latencies, we
first evaluate our approach with synthetic data.
In \cref{sec:covert-align}, the proposed algorithm will be applied to real
\ac{eeg} data.

@@

A sample of the simulated evoked data is displayed in \cref{fig:wcble/sim-sine}.
\fullpagefig{%
  \sffamily
  \sansmath
  \inputpgf{figures/wcble}{simulated-sine.pgf}
}{%
  Simulated target and non-target evoked data.
}{%
  Simulated target (bottom half of epochs) and non-target (top half of epochs) evoked data
  and their average (red: target, yellow: non-target, shaded area: standard
  deviation).
  Targets contain a sine wave pulse jittered by $\sigma$. 32 EEG channels with pink
  spatiotemporal noise were simulated using a forward head model, and channel C3 is
  plotted here.
}{fig:wcble/sim-sine}

@@

\section{Discussion \& conclusion}

>>>>

\chapter{\Acs{erp} latency estimation and alignment}
\label{sec:wcble}
\emph{\Cref{sec:wcble/literature/contrib}, \cref{sec:wcble/methods} (including
\cref{fig:diagram-train,fig:diagram-eval}), and the third-to-last paragraph of
\cref{sec:wcble/conclusion} were published
as part of~\textcite{VanDenKerchove2024}.}

@@

The \ac{wcble} algorithm is presented in \cref{sec:wcble/app/wcble-alg}.
Its training phase is visualized in~\cref{fig:diagram-train}.
The initial training epochs $\{\mat{X}_n^{(1)}\}_{n=1}^N$ are set to $\{\mat{X}_n^\mathrm{train}\}_{n=1}^N$.
At every iteration, classifier $\mathcal{C}$ is trained like in \ac{cble}:
\begin{equation}
  \theta^{(i)} =
  \mathrm{train}_\mathcal{C}(\{\mat{X}^{(i-1)}_n[:,s_1:s_2]\}_{n=1}^N,\mathbf{l}^\mathrm{train})
\end{equation}
Next, latency $s_{l_n}^{(i)}$ is determined for every epoch $\mat{X}^{(i)}$ corresponding
to its class label $l_n$ using \cref{eq:prob-latency}.
Finally, the training epochs $\mat{X}^{(i+1)}$ for the next iteration are determined by aligning
each original training epoch to the latency $s_{l_n}^{(i)}$ corresponding to its respective class
label.
\begin{equation}
  \mat{X}^{(i+1)}_n = \mathrm{align}(\mat{X}_n^\mathrm{train}, s_{l_n}^{(i)}) \quad \forall n=1,\ldots,N
\end{equation}
Aligning is performed by shifting and zero-padding the signal to the right if
the latency is negative relative to the time window onset, and to the left if
positive, by the difference between the latency and the window onset.
\fullpagefig{%
  \includegraphics[width=\textwidth]{figures/wcble/figure1.pdf}
}{%
  Schematic representation of the \acs{wcble} training phase.
}{%
  Schematic representation of the \acf{wcble} training phase.
  (1) The first-stage spatiotemporal binary classifier is trained on a set of
  epochs.
  (2) It is then applied to time-shifted copies of the epochs to obtain scores
  and class probabilities over time.
  (3) The medians of these probability distributions are assumed as the class
  latencies.
  (4) The epochs are aligned to their corresponding class latencies by
  shifting in time such that all latencies fall at the same moment.
  (5) The spatiotemporal classifier is then retrained on the aligned epochs
  for the next iteration.
  (6) After the iterative process halts, the scores and latencies obtained
  from the last iteration are used to train the second-stage
  classifier.
}{fig:diagram-train}
The process halts after a fixed number of iterations or when the estimated set
of latencies has been encountered before, indicating it ended up in a loop.
In the end, the procedure should result in enhanced classifier parameters $\theta^*$,
closer to those when there would be no jitter between epochs.
Note that using the median approach for robust latency estimation results in a
smoother yet longer convergence process compared to the maximum/minimum approach.
We can then apply the classifier with enhanced parameters $\theta^*$ in a \ac{cble}
manner to unseen epochs as illustrated in \cref{fig:diagram-eval} to obtain a
vector of scores over time as in \cref{sec:wcble/methods/cble} and the estimated latencies as
in~\cref{sec:robust-latency}.
\fullpagefig{
  \includegraphics[width=\textwidth]{figures/wcble/figure2.pdf}

}{%
  Schematic representation of the \acs{wcble} test phase.
}{%
  Schematic representation of the \acf{wcble}. (1) The
	first-stage spatiotemporal binary classifier obtained from the training phase is
	applied to time-shifted copies of the epochs to obtain scores and class probabilities
	over time. (2) The medians of these probability distributions are assumed
	as the new class latencies. (3) The scores and class latencies are input to the
  trained second-stage classifier, which predicts the label of the epochs.
}{fig:diagram-eval}
\section{Results}
\label{sec:wcble/results}
Since it is hard to obtain ground truth measures of \ac{erp} latencies, we
first evaluate our approach with synthetic data.
In \cref{sec:covert-align}, the proposed algorithm will be applied to real
\ac{eeg} data.

@@

A sample of the simulated evoked data is displayed in \cref{fig:wcble/sim-sine}.
\fullpagefig{%
  \sffamily
  \sansmath
  \inputpgf{figures/wcble}{simulated-sine.pgf}
}{%
  Simulated target and non-target evoked data.
}{%
  Simulated target (bottom half of epochs) and non-target (top half of epochs) evoked data
  and their average (red: target, yellow: non-target, shaded area: standard
  deviation).
  Targets contain a sine wave pulse jittered by $\sigma$. 32 \ac{eeg} channels with pink
  spatiotemporal noise were simulated using a forward head model, and channel C3 is
  plotted here.
}{fig:wcble/sim-sine}

@@

\section{Discussion \& conclusion}
\label{sec:wcble/conclusion}
